---
title: "Binomial"
author: "Nick Wendt"
date: "2025-02-13"
output: html_document
---

### Calculates the exact probability of x successes

$p(X=x)\ where\ X \sim Bin(n,\pi)$

```{r}
prob <- dbinom(5, size = 10, prob = 0.5)
calc.prob <- choose(10,5) * 0.5^10 
```

### Calculates the cumulative probability

$p(X\leq x)\ where\ X \sim Bin(n,\pi)$

```{r}
prob <- pbinom(5, size = 10, prob = 0.5)
l <- c(0,1,2,3,4,5)
calc.prob <- sum(choose(10,l)*0.5^10)
```

### Generates n random binomial values

```{r}
observations <- rbinom(20, 50, 0.5)
# Generates 20 random binomial values
# Each value represents the number of successes
# Probability of success is 0.5
# Number of trials is 50
print(observations)
```

### Plotting the Binomial Distribution

```{r}
plot(0:10, dbinom(0:10, 10, 0.2), type = "h",
     xlab = "x", ylab = "P(x)")
# Plots a binomial distribtution with n = 10
# number of trials goes 0 - 10
```

### Finding the maximum likelihood curve

```{r}
bin <- curve(dbinom(2,3,x), xlim = c(0,1))
# First number in dbinom is number of successes
# Second number is number of trials
# x is the probability of success
# this dbinom is seeing 2 successes as a function of x
# where x is the true value of p
```

### Score test for proportions

```{r}
prop.test(837, 1810, p=0.50, alternative="two.sided", correct=FALSE)
```

The test statistic calculated in the code above is as follows:

$$
Z = \frac{\hat{\pi}-\pi_0}{\sqrt{\frac{\pi_0(1-\pi_0)}{n}}}
$$

Where: $\hat{\pi} = \frac{\sum x_i}{n}$ which is the MLE

and $\pi_0$ is the proposed value of $\pi$ under the null hypothesis

in this example $\hat{\pi} = 837/1810 \approx 0.4624$ and $\pi_0  =0.5$

Calculating the test statistic

```{r}
z <- ((837/1810)-0.5)/sqrt((0.5*0.5)/1810)
p.val <- 2 * (1 - pnorm(abs(z)))
p.val
```

### Wald Test

The only difference for the Wald test is that we use $\hat{\pi}$ to calculate the standard error rather than $\pi_0$

$$
Z = \frac{\hat{\pi}-\pi_0}{\sqrt{\frac{\hat{\pi}(1-\hat{\pi})}{n}}}
$$

```{r}
p.hat <- 837/1810
wald.z <- (p.hat-0.5)/sqrt((p.hat*(1-p.hat))/1810)
wald.p.val <- 2 * (1- pnorm(abs(wald.z)))
print(wald.p.val)
```

Note:

Use Wald test for large sample sizes and when $\hat{\pi}$ is not close to 0 or 1

Use score test for medium sample sizes

### Confidence Interval

$$
\hat{\pi} \pm Z_{\alpha/2} \times SE
$$

Where $SE=\sqrt{\frac{\hat{\pi}(1-\hat{\pi})}{n}}$ for Wald and $SE=\sqrt{\frac{\pi_0(1-\pi_0)}{n}}$ for Score
