---
title: "Nonlinear Models"
author: "Nick Wendt"
date: "2025-04-03"
output: html_document
---

## Types of Models

Linear in variables ( $X_{i's}$ ) and parameters ( $\beta_{i's}$ ) :

$$
Y+\beta_0+\beta_1X_1+\beta_4X_4+\epsilon
$$

Linear in parameters, nonlinear in variables:

$$
\begin{align*}Y &= \beta_0 + \beta_1 X_1^2 + \beta_2 \log X_2 + \beta_3 \sqrt{X_3} + \epsilon \\Y &= \beta_0 + \beta_1 Z_1 + \beta_2 Z_2 + \beta_3 Z_3 \\\text{where} \quad Z_1 &= X_1^2,\quad Z_2 = \log X_2,\quad Z_3 = \sqrt{X_3}\end{align*}
$$

Nonlinear in parameters:

$$
Y=\beta_0 +\beta_1X_1+\beta_2X_2^{\beta_4}+\beta_5\beta_6X_4+\epsilon
$$

## When to use Nonlinear Models

If two variables have an exponential relationship, then an exponential growth model could be

$$
Y=\beta_0e^{\beta_1X}+\epsilon
$$

```{r}
X <- seq(0, 10, by=0.5)  # Time in years
Y <- 100 * exp(0.3 * X) + rnorm(length(X), mean=0, sd=50) 
# rnorm is to simulate random noise (error)  

# Transform Y for linear regression: Take the natural log
# This IS A LINK FUNCTION
log_Y <- log(Y)

# Fit a linear model on log-transformed data
exp_model <- lm(log_Y ~ X)


plot(X, Y, main="Exponential Relationship Example",
     xlab="Time (Years)", ylab="Population")


X_seq <- seq(min(X), max(X), length.out=100)
Y_pred <- exp(predict(exp_model, newdata = data.frame(X = X_seq)))
lines(X_seq, Y_pred, col="red", lwd=2)
```

Or if two variables have a logarithmic relationship:

$$
Y=\beta_0+\beta_1log(X)+\epsilon
$$

```{r}
X <- seq(1,100, by = 1)
Y <- 10 + 5 * log(X) + rnorm(100, mean = 0, sd = 1) 

log.model <- lm(Y ~ log(X))

plot(X, Y, main = "Log Relationship Example", xlab = "Advertising Budget", ylab = "Sales Revenue")
curve(coef(log.model)[1] + coef(log.model)[2] * log(x), 
      from=min(X), to=max(X), col="red", lwd=2, add=TRUE)
```

## Loglinear Models

$$
Y= \beta_0e^{\beta_1X}
$$

This model is non-linear in parameters and variables

If we take the derivative of this model we get:

$$
\frac{dY}{dX}=\beta_0e^{\beta_1X}\beta_1
$$

If we instead take the log of this model, $log(\beta_0 e^{\beta_1X})$ we get $log(\beta_0)+\beta_1X$ which is a linear model. $log(Y)=log(\beta_0)+\beta_1X$ leads to the interpretation of $\beta_1 \times 100 \approx semielasticity$

Rule of thumb: use % for variables that appear in log() form.

In a logarithmic model $Y=\beta_0X^{\beta_1} \rightarrow log(Y)=log(\beta_0)+\beta_1 log(X)$ where $\hat{\beta_1}$ can be interpreted as the elasticity between X and Y. "If there is a 1% change in X, then there is a $\hat{\beta_1}\%$ change in Y.

## Quadratic Models

Model: $Y=\beta_0+\beta_1X_1+\beta_2X_1^2$

The effect of $X_1$ can vary depending on the value of $X_1$ . This can be used to represent diminishing marginal product in a production function

$\beta_1$ can be interpreted as the effect of $X_1$ on Y only when $X_1=0$ (often a weird interpretation)

### Interaction Terms

Given a model:

$$
Y=\beta_0+\beta_1X_1+\beta_2X_2+\beta_3 X_1X_2
$$

The effect of X on Y:

$$
\frac{dY}{dX_1} = \beta_1+\beta_3X_2 \\
\frac{dY}{dX_2}=\beta_2+\beta_3X_1
$$

Effect of $X_1$ can depend on $X_2$ and vice versa

$\beta_3$ may indicate synergies or antagonism between the effects of regressors
