---
title: "Proportional Hazard"
author: "Nicholas Wendt"
date: "2025-03-10"
output: pdf_document
---

## What is a Proportional Hazards Model

A proportional hazards model is a class of survivor models. These models show the amount of time passing before an event occurs. The hazard rate at time $t$ is the probability per short time $dt$ that an event will occur between time $t$ and $t +dt$ given that no event has occurred before time $t$ .

### Prop Hazard Model:

$$
h(t|X)=h_0(t)exp(\beta_1X_1+\beta_2X_2+...+\beta_pX_p)
$$

-   $h_0(t)$ is the baseline hazard function (the hazard when all predictors are 0)

-   $X_1,X_2, ..., X_p$ are the predictor variables

-   $\beta_1,\beta_2,...,\beta_p$ are the regression coefficients

### Assumptions

The key assumption is that the hazard ratio (relative risk) between two individuals remains constant over time. This means that if $p_1$ has twice the risk as $p_2$ , they will always have twice the risk at any other time.

```{r}
library(survival)
data(lung)
head(lung)
```

```{r}
surv_obj <- Surv(time = lung$time, event = lung$status)
cox_model <- coxph(surv_obj ~ age + sex + ph.ecog, data = lung)
summary(cox_model)
```
