---
title: "hw_5"
author: "Nicholas Wendt"
date: "2025-03-11"
output: pdf
---

Questions: 2.18, 2.21, 2.22, 2.28

## Question 2.18

Yes, it is valid to apply the chi-squared test of independence here since all counts in each cell are ≥ 5 In this table, each factor (A, B, C) represents a separate categorical variable. We can use this table to create 3 separate 2x2 contingency tables. Since the number of men and women who disagree with each statement are implied (100-number who agree), we have all the information we need to cross-classify gender with each of the three statements.

```{r}
contingency.table <- data.frame(Gender = c("Men", "Women"),
                                Agree_A = c(60,70),
                                Disagree_A = c(40,25))
print(contingency.table)
```

## Question 2.21

#### Part a

```{r}
satisfaction.table <- matrix(data = c(2,4,13,3,2,6,22,4,0,1,15,8,0,3,13,8), byrow = TRUE, ncol = 4, nrow = 4)
colnames(satisfaction.table) <- c(">25","25-15","15-5","5>")
rownames(satisfaction.table) <- c("very satisfied","moderately satisfied", "little satisfied","very dissatisfied")

print(satisfaction.table)
```

```{r}
satisfaction.chi <- chisq.test(satisfaction.table)
print(satisfaction.chi)
print(satisfaction.chi$residuals)
```

The $\chi^2$ test for independence works poorly on data that has cell counts \< 5. Since our data has cells with low counts, the $\chi^2$ test statistic maybe be inaccurate.

$$
\chi^2 = \sum_{\forall ij}\frac{(O_{ij}-E_{ij})^2}{E_{ij}}
$$

If we choose to believe this $\chi^2$ statistic, we would fail to reject since:

$$
p-val = 0.242 > \alpha = 0.05
$$

and conclude that job satisfaction and salary are independent.

#### Part b

```{r}
fisher.test(satisfaction.table)
```

## Question 2.22
