---
title: "HW6"
author: "Nicholas Wendt"
date: "2025-04-08"
output: pdf_document
---

Questions: 3.3, 3.5, 3.8

## Question 3.3

#### Part a

The prediction equation is: $E(Y)=0.00255+0.00109x$

Where x = mother's alcohol consumption and Y = whether a baby has sex organ malformation

The intercept is the probability of a baby having a sex organ malformation when the mother doesn't consume alcohol during pregnancy

The slope is the increase in probability of a baby having a sex organ malformation due to a 1 unit increase in alcohol consumption

#### Part b

(i) 

```{r}
model.0 <- 0.00255+0.00109*0
model.7 <- 0.00255+0.00109*7
print(model.0)
print(model.7)
```

(ii) 

```{r}
rr <- model.7/model.0
rr
```

The a baby is about 4 times more likely to have a sex organ malformation when the mother consumes 7 units of alcohol compared to a baby who's mother consumed no alcohol.

#### Part c

```{r}
alcohol.consumption <- matrix(data = c(17066,48,14464,38,788,5,126,1,37,1), byrow = TRUE, nrow = 5, ncol = 2)
colnames(alcohol.consumption) <- c("Absent", "Present")
rownames(alcohol.consumption) <- c("0","<1","1-2","3-5",">5")
alcohol.consumption
new.scores <- c(0,1,2,3,4)
old.scores <- c(0,0.5,1.5,4,7)
alcohol1=data.frame(alcohol.consumption,old.scores, new.scores)
alcohol1
```

```{r}
alc.model.new = glm( cbind(Present, Absent) ~ new.scores,family=binomial(link = "identity"), data=alcohol1 )
alc.model.old <- glm( cbind(Present, Absent) ~ old.scores,family=binomial(link = "identity"), data=alcohol1)
summary(alc.model.new)
summary(alc.model.old)
```

```{r}
model.0 <- 0.0025977+0.0005044*0
model.7 <- 0.0025977+0.0005044*7
rr <- model.7/model.0
rr

(fit_old=fitted(alc.model.old))
(fit_new=fitted(alc.model.new))

rr_old=fit_old[5]/fit_old[1]
rr_new=fit_new[5]/fit_old[1]
rr_old
rr_new
```

#### Part d

```{r}
altered.alc <- matrix(data = c(17066,48,14464,38,788,5,126,1), byrow = TRUE, nrow = 4, ncol = 2)
colnames(altered.alc) <- c("Absent", "Present")
rownames(altered.alc) <- c("0","<1","1-2","3-5")

scores <- c(0,0.5,1.5,4)
alter.w.scores <- data.frame(altered.alc, scores)
altered.model = glm( cbind(Present, Absent) ~ scores,family=binomial(link = "identity"), data=alter.w.scores )
summary(altered.model)
```

```{r}
altered.0 <- 0.0026026 + 0.0008184*0
altered.7 <- 0.0026026 + 0.0008184*7
rr <- altered.7/altered.0
rr
```

The model is sensitive to this change.

## Question 3.5

```{r}
crab <- read.table("https://users.stat.ufl.edu/~aa/cat/data/Crabs.dat", header = TRUE)
crab
```

#### Part a

```{r}
crab.model <- glm(y ~ weight, data = crab, family = binomial(link = "identity"), start = c(0.01, 0.001))
summary(crab.model)
```

```{r}
predict(crab.model, newdata = data.frame(weight = 5.2))
predict
```

The linear model predicts that a crab with a width of 5.2 will have a satellite

#### Part b

```{r}
crab.log <- glm(y ~ weight, data = crab, family = binomial)
summary(crab.log)
```

```{r}
predict(crab.log, newdata = data.frame(weight = 5.2), type = "response")
```

the logistic regression predicts that there is a %99.68% chance for a crab with width 5.2 to have a satellite.

## Question 3.8

```{r}
snoring <- matrix(c(24,1355,35,603,21,192,30,224), nrow = 4, ncol = 2, byrow = TRUE)
colnames(snoring) <- c("Yes","No")
rownames(snoring) <- c("Never", "Occasional", "Nearly every night", "Every night")
snoring
```

```{r}
scores <- c(0,2,4,5)
snoring.df <- data.frame(snoring, scores)
snoring.model <- glm(cbind(Yes,No) ~ scores, data = snoring.df, family = binomial(link = "logit"))
summary(snoring.model)
```

$H_0: \beta_1 = 0$

$H_a: \beta_1 \neq 0$

Reject $H_0$ since p-val = $1.94e-15 < \alpha = 0.05$

Snoring is statistically significant when predicting heart disease

```{r}
confint(snoring.model)
```

Since 0 isn't in the CI we can reject.
