---
title: "HW7"
author: "Nicholas Wendt"
date: "2025-04-16"
output: pdf_document
---

Questions: 3.11, 4.1, 4.5

## Question 3.11

```{r}
treat.a <- c(8, 7, 6, 6, 3, 4, 7, 2, 3, 4)
treat.b <- c(9, 9, 8, 14, 8, 13, 11, 5, 7, 6)


treatment <- c(rep("A", 10), rep("B", 10))

#X = 1 for treatment b
#X = 0 for treatment a

x <- ifelse(treatment == "B", 1, 0)
treat.total <- c(treat.a, treat.b)

treat.data <- data.frame(treat.total, x)

treat.model <- glm(treat.total ~ x, family = poisson, data = treat.data)
summary(treat.model)
exp(coef(treat.model))
```

The prediction equation is $number\ of\ imperfections = 1.6094 + 0.5878 x$ where $x=1$ for treatment b and $x=0$ for treatment a. $\hat{\beta} = e^{0.5878} \approx 1.8$ which means the rate of imperfections under treatment b is around 1.8 times higher than the rate of imperfections for treatment a. On average, wafers treated with treatment b will have 80% more imperfections compared to wafers treated with treatment a.

## Question 4.1

```{r}
LI <- c(8,8,10,10,12,12,12,14,14,14,16,16,16,18,20,20,20,22,22,24,26,28,32,34,38,38,38)
y <- c(0,0,0,0,0,0,0,0,0,0,0,0,0,1,1,1,0,1,0,0,1,1,0,1,1,1,0)

cancer.data <- data.frame(y,LI)
colnames(cancer.data) <- c("y", "LI")
cancer.data
```

#### Part a

```{r}
cancer.model <- glm(y ~ LI, family = binomial, data = cancer.data)
predict(cancer.model, newdata = data.frame(LI = 26), type = "response")
```

#### Part b

```{r}
exp(coef(cancer.model)[2])

```

#### Part c

$\hat{P}(Y=1)$ changes over the range at a rate of $\beta \pi(x)(1-\pi(x))$ . At the median effect level, when $x=-\alpha/\beta$ , the rate of change is $0.25\beta$ .

```{r}
quantile(cancer.data$LI, probs = c(0.25, 0.75), na.rm = TRUE)
```

```{r}
p25 <- predict(cancer.model, newdata = data.frame(LI = 13), type = "response")

p75 <- predict(cancer.model, newdata = data.frame(LI = 25), type = "response")

change <- p75 - p25
change
```

This means that the difference in probability between 13% labeled cells achieving remission and 25% labeled cells achieving remission is 0.33

#### Part d

```{r}
beta <- coef(cancer.model)[2]
estimate.y <- predict(cancer.model, newdata = data.frame(LI = 8),type = "response")
rate.of.change <- beta*estimate.y*(1-estimate.y)
(rate.of.change)
```

#### Part e

```{r}
beta <- coef(cancer.model)["LI"]

p.hat <- predict(cancer.model, type = "response")
marginal.effects <- beta * p.hat * (1 - p.hat)

average.marginal.effect <- mean(marginal.effects)
average.marginal.effect
```

The estimated average marginal effect is about 0.0226 meaning that a 1 unit increase in LI leads to a 2.26% increase in probability that Y=1. Points closer to the middle will have a marginal effect larger than 2.26% while points closer to the ends will have a marginal effect less than 2.26%.

## Question 4.5
