---
title: "Introduction to R"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
```

## Entering data

### variables

```{r}
x=3  
y=4
3*x
x+y
y^2
```

### Using c() function

```{r}
my.data=c(11,14,22,15)
z <- c(5,7,8,12)
#this code will create your dataset and assign it the variable name my.data
my.data    #If you want to know what a variable holds, you must ask for it by name
my.data+500
2*my.data
length(my.data)

#using :
stats.data=1:5   #this code will create your sequence of data from 1 to 5, by increments of 1
stats.data   

decreasing.data=5:-6   #this code will create your sequence of data from 5 to -6, by increments of -1
decreasing.data   

#using seq() function
seq(0,3,by=2)   # seq(low.number, high.number, increment.size)
seq(0,3,length=5)  # seq(low.number, high.number, sequence.length)

#using rep() function
rep(c(1,2,3), each=3) 
rep(c(1,2,3), times=3) 
```

## Math Functions

```{r}
2 + 4 * 5      # Order of operations
log (10)       # Natural logarithm with base e=2.7182
log10(5)      # Common logarithm with base 10
5^2             # 5 raised to the second power
5/8             # Division
sqrt (16)      # Square root
abs (3-7)     # Absolute value 
pi                # 3.14
exp(2)         # Exponential function 
round(pi,0)  # Round pi to a whole number
round(pi,1)  # Round pi to 1 decimal place
round(pi,4)  # Round pi to 4 decimal places
floor(15.9)   # Rounds down
ceiling(15.1)  # Rounds up
cos(.5)        # Cosine Function
sin(.5)        # Sine Function
tan(.5)       # Tangent Function
acos(0.8775826)        # Inverse Cosine
asin(0.4794255)       # Inverse Sine
atan(0.5463025)      # Inverse Tangent
```

##Data Manipulations

```{r}
x=c(1,3,2,10,5)    #create a vector x with 5 components 
### Sort your data
sort(x)                # increasing order 
sort(x, decreasing=T)  # decreasing order 
```

Component extraction is a very important part of vector calculation.

```{r}
x 
length(x)           # number of elements in x 
x[4]                # the fourth element of x 
x[2:5]              # the second to fifth element of x, inclusive 
x[-2]               # all except the second element 
x[-c(1,2)]          # all except the first and the second elements
x[x>3]              # list of elements in x greater than 3 
```

Logical vector can be handy:

```{r}
x>3                   # Which x values are greater than 3
sum(x>3)            # number of elements in x greater than 3 
```

##Numerical Summaries

```{r}
x=c(34, 24, 10, 16, 52, 76, 33, 31, 46, 24, 18, 26, 57, 32, 25, 48, 22, 48, 29, 19)
x
max(x)
min(x)
which.min(x)
which.max(x)
mean(x)
mean(x, trim = 0.1)
median(x)
sd(x)
var(x)
range(x)
sum(x)
summary(x) # Calculate five-number summary: min,Q1,median,Q3,max, return an object
my.q=fivenum(x) # calculate five-number summary, return a numeric vector
quantile(x, c(.25, .5, .75, 1)) ### Calculates the numbers associated to defined percentiles
quantile(x,0.99)
iqr=IQR(x)

q1=my.q[2]
q3=my.q[4]
q1-1.5*iqr    #lower fence for outliers
q3+1.5*iqr    #upper fence for outliers
q1-3*iqr      #lower fence for extreme outliers
q3+3*iqr      #upper fence for extreme outliers
```

### Built-in data frames

We need to load a library to work with the battery dataset

```{r}
#install.packages("EngrExpt")
library(EngrExpt);
data(battery); 
attach(battery)

```

Let's view several lines of actual dataset

```{r}
head(battery)
```

The following 3 lines of code will create three separate datasets by battery type

```{r}
type.a=subset(battery, type=="A")

type.b=subset(battery, type=="B")

type.c=subset(battery, type=="C")
```

Let's take a look at dataset type.a

```{r}
head(type.a)
```

Let's look at just the lifetime variable within the type.a dataset

```{r}
type.a$lifetime      ### DataSetName$VariableName
```

Now, we can analyze the lifetime variable within the type.a dataset

```{r}
mean(type.a$lifetime)
sd(type.a$lifetime)
```

more built-in dataframes <https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html>

##Graphical Categorical

EXAMPLE OF THE table FUNCTION

```{r}
x=c("Yes", "No", "Yes", "Yes", "No")  # Create categorical data
x

table(x)  # table function returns the frequency for each category of data
```

Now we will use storm/hurricane data

```{r}
sh = load("/inclass_activities/data/u.s._hurricane_data2011.txt")
cat = sh$cat
year = sh$year
```

Create a table for the category of storm/hurricane

```{r}
cat.counts=table(cat)
```

Create a table for the year of the storm/hurricane

```{r}
year.counts=table(year) 
```

# View the respective tables

```{r}
cat.counts

year.counts  
```

## CREATING BAR GRAPHS

### Create a barplot of storm/hurricane categories

```{r}
barplot(cat.counts, col="red") 
```

###Create a barplot using percentages for the years

```{r}
barplot(year.counts/length(year), col="green") 
```

## CREATING PIE CHARTS

###Basic pie chart

```{r}
pie(cat.counts) 
```

# Add names to your piechart

```{r}
names(cat.counts)=c("2005", "2006", "2007", "2008", "2009", "2010")  

# Pie chart with names printed out
pie(cat.counts) 

# Now, add a little color and a main :)  
pie(cat.counts, col=rainbow(12), main="Atlantic Hurricanes by Year") 
```

## Multivariate data

```{r}
library(MASS)
data(UScereal)
head(UScereal)
summary(UScereal)

UScereal$shelf=as.factor(UScereal$shelf)
summary(UScereal)

attach(UScereal)
hist(sodium) 
abline(v=mean(sodium),lty=3)
plot(potassium, protein) 

boxplot(sugars) 
```
